% LaTeX file for resume 
% This file uses the resume document class (res.cls)
\RequirePackage{cmap}
\documentclass{res} 
%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
%\usepackage{newcent}   % uses new century schoolbook postscript font 
\setlength{\textheight}{9.5in} % increase text height to fit on 1-page 

% Colored links avoid PDF link rectangles that sit on top of the header rule.
\usepackage[colorlinks=true,urlcolor=blue,linkcolor=black]{hyperref}
\usepackage{fancyhdr}
% T2A + T1: Cyrillic (Russian) and Latin; last listed encoding (T1) is the default.
\usepackage[T2A,T1]{fontenc}
\usepackage{lmodern}
\usepackage{cmap}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
% Cyrillic (UTF-8): use \foreignlanguage{russian}{...} or
% {\selectlanguage{russian}...\selectlanguage{english}} so LaTeX switches to T2A fonts.
\usepackage{needspace}

% Tighter page margins; widen \resumewidth so text still fits between margins.
% Narrower section-title column (default overlapped style was 0.5in).
\makeatletter
\addtolength{\topmargin}{-30pt}
\addtolength{\textheight}{30pt}
\addtolength{\oddsidemargin}{-24pt}
\addtolength{\evensidemargin}{-24pt}
\advance\resumewidth by 48pt
\newsectionwidth{0.33in}
\setlength{\headwidth}{\resumewidth}
% fancyhdr: \fancyheadoffset[OL]{L} sets head width to \textwidth+L (here L=\sectionwidth
% = \resumewidth) and inserts \hskip-L so the header moves left by \sectionwidth, matching
% overlapped \section titles (res.cls sets \hoffset=\sectionwidth).
\fancyheadoffset[OL]{\sectionwidth}
% res.cls default is 3.5ex plus 1ex minus .2ex before each \section.
\sectionskip=2ex plus 0.35ex minus 0.12ex
\makeatother

% Running header (res.cls sets \headheight and \headsep to 0).
\setlength{\headheight}{29pt}
\setlength{\headsep}{4pt}
\addtolength{\textheight}{-\dimexpr\headheight+\headsep\relax}
\newcommand{\ResumeHeaderLeft}{\Large\bfseries Sophia Kovaleva}
\newcommand{\ResumeHeaderCenter}{Privacy, AI/ML, data science}
\newcommand{\ResumeHeaderRight}{\href{mailto:sophia.m.kovaleva@gmail.com}{sophia.m.kovaleva@gmail.com}}
\fancyhf{}
\fancyhead[L]{\ResumeHeaderLeft}
\fancyhead[C]{\ResumeHeaderCenter}
\fancyhead[R]{\ResumeHeaderRight}
\renewcommand{\headrulewidth}{0.4pt}
% Double rule across full \headwidth (= \resumewidth, i.e.\ body + section gutter).
\renewcommand{\headrule}{%
  \begingroup
  \nointerlineskip
  \hrule height \headrulewidth width \headwidth\relax
  \kern 0.9pt
  \hrule height \headrulewidth width \headwidth\relax
  \endgroup
}
\fancyfoot{}
\pagestyle{fancy}
\fancypagestyle{plain}{%
  \fancyhf{}%
  \fancyhead[L]{\ResumeHeaderLeft}%
  \fancyhead[C]{\ResumeHeaderCenter}%
  \fancyhead[R]{\ResumeHeaderRight}%
  \renewcommand{\headrulewidth}{0.4pt}%
  \fancyfoot{}%
}

% One-line job header: dates left, title centered, company/location right; rule below.
% Keep header with the following text/list (\Needspace + \nopagebreak).
\newcommand{\ExperienceHeader}[3]{%
  \Needspace{9\baselineskip}%
  \par\vspace{0.35ex}%
  \noindent
  \begin{tabular*}{\textwidth}{@{}l@{\extracolsep{\fill}}c@{\extracolsep{\fill}}r@{}}
    #1 & \textbf{#2} & #3 \\[-0.2ex]
    \hline
  \end{tabular*}%
  \par\nopagebreak[4]\vspace{0.35ex}\nopagebreak[4]%
}

\begin{document} 
\begin{resume}
\vspace{-20pt}% gap below header / after opening; res.cls uses \parskip between blocks

% Match left edge of overlapped \section titles (body column minus \sectionwidth).
\noindent\hspace{-\sectionwidth}%
\begin{tabular*}{\dimexpr\textwidth+\sectionwidth\relax}{@{}l@{\extracolsep{\fill}}r@{}}
	Mountain View, CA & Phone: \href{tel:+13024154176}{(302) 415-4176} \\
	GitHub: \href{https://github.com/kosonya/}{github.com/kosonya} &
	LinkedIn: \href{https://www.linkedin.com/in/sophia-kovaleva-1a664a3b/}{linkedin.com/in/sophia-kovaleva-1a664a3b} \\
\end{tabular*}%

\section{SUMMARY}

Senior privacy engineer, ML researcher, data scientist, with 15+ years of experience in research and development. Left CMU PhD program to focus on high-impact industry work. Looking for opportunities to apply my diverse skills in a high-impact role.

\section{EDUCATION}
\begin{itemize}
	\item
		\underline{Carnegie Mellon University, July 2013 -- February 2019}
		\begin{itemize}
			\item
				Master of Science in Electrical and Computer Engineering
			\item
				PhD (All But Dissertation, unfinished). Advisor: Dr. Anupam Datta, PhD thesis: ``Representation-aware accounting of machine learning models''.
		\end{itemize}
	\item
		National Research University HSE MIEM, September 2007 -- June 2012\\
		Bachelor of Science (Engineer) in Computer Science
\end{itemize}

\section{SKILLS}          
\begin{itemize}
	\item
		Primary programming language: Python.
	\item
		Secondary: Java, SQL, C, C++, JavaScript, Haskell, Fortran, \LaTeX{}.
	\item
		Building ML systems and data pipelines with TensorFlow, Apache Spark, scikit-learn, NumPy, Jupyter, Django, PostgreSQL, Redis.
	\item
		Data visualization with matplotlib, GNU Octave, MATLAB, D3.js.
	\item
		Analysis of privacy risks on mobile devices, especially against local attacks against at-risk users.
	\item
		Data anonymization, differential privacy, analysis of code, logs, and technical documentation for privacy risks.
	\item
		Embedded systems (AVR, ARM), wireless communication, geolocation, sensor fusion, and mobile device security.
	\item
		General knowledge of GNU/Linux, networking, Docker, agentic AI, and LLMs. General interest in molecular biology and neuroscience.
\end{itemize}
 
\section{EXPERIENCE}

\ExperienceHeader{Spring 2025 -- now}{}{Sabbatical, exploring early startup ideas}

\ExperienceHeader{Spring 2019 -- Spring 2025}{Senior Privacy Engineer}{Google, Mountain View, CA}
\begin{itemize}
	\item
		I conducted pre-launch privacy reviews for Android features being rolled out to over 3 billion users.
	\item
		I led the effort to modify the accelerometer API to mitigate the ``Gyrophone'' eavesdropping attack.
	\item
		I worked with product teams and led long-term privacy improvements in on-device intelligence, federated analytics, at-risk users privacy, and protection against coercive attacks.
	\item
		I drove improvements in the privacy policies and practices for end-to-end encryption, Nearby Share, preloaded apps, notice and consent.
	\item
		I built internal data analytics pipelines to identify problematic data collection and use practices, cross-link them to the responsible pieces of code, and work with the teams to fix them.
	\item
		I was the primary reviewer for the Android Vulnerability Reward Program reports. I built data analytics pipelines to identify inefficiencies in the internal review processes and led the effort to amend them accordingly.
	\item
		I started and drove long-term R\&D initiatives in the areas of E2EE voice calling and better key management systems for file-based encryption.
	\item
		I managed one direct report.
\end{itemize}

\ExperienceHeader{Fall 2013 -- Spring 2019}{PhD Student / Research Assistant}{Carnegie Mellon University, Moffett Field, CA}
\begin{itemize}
	\item
		Improving Wireless Emergency Alerts -- I built a data pipeline from Twitter to do sentiment analysis of people's reactions to WEAs. I developed a mobile app to collect data in the field and test various improved geofencing techniques.
	\item
		Room Occupancy Detection in SmartSpace -- I used data from wall-mounted physical sensors to estimate the presence and the number of people in rooms.
	\item
		Wireless communication set-up for RSSI broadcast -- I estimated 3D radiation pattens of cellular antennae by interpolating RSSI data collected by a drone-based platform.
	\item
		Considerate systems -- I worked on a system for estimating the speaking patterns in video conference calls and nudging the participants to be more considerate of each other.
	\item
		Explaining recommender systems -- I built a matrix factorization recommender system with public Netflix data and a supervised learning model to explain the recommendations.
\end{itemize}

\ExperienceHeader{Summer 2015}{Data Science Intern}{Zugata, Inc., Palo Alto, CA}
\begin{itemize}
	\item
		I was the only data scientist on the team of 5 people. I took leadership in developing the strategy and direction of analytics and interviewed candidates for the data science role.
	\item
		I built an end-to-end pipeline for collecting metadata about internal corporate communications, building a social graph of the company, and making suggestions for asking people for peer feedback.
	\item
		I used PostgreSQL, Redis, Python, Django, Nginx, and D3.js to collect the communications, analyze them, serve peer feedback nudges through a REST API, and visualize the social graph in a web application.
\end{itemize}


\ExperienceHeader{Fall 2010 -- June 2013}{Technician / TA}{MIEM / MIEM HSE, Moscow, Russia}
\begin{itemize}
	\item
		As a sophomore, I started volunteering to develop a curriculum and started teaching some of the lectures for a programming class to freshmen. By my third year, when I was formally hired, I was largely responsible for most of the teaching and curriculum.
	\item
		I developed curriculum for the intro C programming class.\\
		(see \href{https://web.archive.org/web/20130414102421mp_/http://wiki.auditory.ru/%D0%9A%D0%B0%D1%82%D0%B5%D0%B3%D0%BE%D1%80%D0%B8%D1%8F:%D0%AF%D0%B7%D1%8B%D0%BA_%D0%A1%D0%B8}{https://web.archive.org/web/20130414102421mp\_/http://wiki.auditory.ru/Категория:Язык\_Си}).
	\item
		I later switched the curriculum to Python and programming 2D games in pygame, and then switching to C and SDL in the next semester.
	\item
		I developed and taught an applied programming class to sophomores. Language: Python. Topics: Python object model, basic functional programming, multiprocessing, socket network interface, GTK-based GUIs, databases and others.
	\item
		I started my research project: ML-enabled tracking of public transit vehicles. I built an end-to-end pipeline for collecting the data with and SBC, storing it in a database, training an ML model on it, and displaying the data in an Android app. I received multiple awards and research grants for my work, even as an undergrad.
\end{itemize}
 
\section{PUBLICATIONS}
\begin{itemize}
	\item
		Datta, A. et al (2018)  Latent Factor Interpretations for Collaborative Filtering, \url{https://arxiv.org/abs/1711.10816}.
	\item
		Erdogmus, H. et al (2015) Opportunities, Options, and Enhancements for the Wireless Emergency Alerting Service, Department of Homeland Security, First Responders Group, February 2016.  
	\item
		Kovalev, M. (2014) Indoor Positioning of Mobile Devices by Combined Wi-Fi and GPS Signals. \emph{Indoor Positioning and Indoor Navigation (IPIN), 2014 International Conference on}, pp.332-339, 27-30 Oct. 2014, doi: 10.1109/IPIN.2014.7275500
	\item
		Kovalev, M. (2013). Testing and the analysis of the performance of an algorithm for positioning public transportation vehicles by GSM signal. \emph{Proceedings of the annual scientific and technical conference of undergraduate and graduate students, and young experts of the Moscow State Institute of Electronics and Mathematics, National Research University Higher School of Economics} (pp. 140--141). Moscow, Russia: MIEM NRU HRE Press. ISBN 978-5-94768-066-9.
\end{itemize}


\end{resume}
\end{document}
